### 2021-10-24 paper/pldi22/formal.tex
Till semantics of instructions

\[
  \begin{array}{lr@{~}c@{~}c@{~}c@{~}l}
    \text{Functions} & \funcset &\ni& \func &::=&
    \kwdef \; x \; \kwrl x^* \kwrr \; \kwcl [\lab: \inst]^* \kwcr\\

    \text{Variables} & \varset &\ni& x\\

@@

    \text{Instructions} & \instset &\ni& \inst &::=&
    \refer = \expr \mid
    x = \expr \kwrl \expr^* \kwrr \mid
    \kwif \; \expr \; \lab \; \lab \mid
    \kwret \; \expr\\

@@

    \text{References} & \referset &\ni& \refer &::=&
    x \mid \refer \kwsl \expr \kwsr\\
  \end{array}
\]

\todo update explanation

A function $\func \in \funcset$ is defined with its name, parameters, and body
instructions with labels $\labset$.  An instruction $\inst \in \instset$ is
a primitive value assignment $x = \pval$,
an operation between values $x = \op \kwrl x^* \kwrr$,
a function call $x = x \kwrl x^* \kwrr$,
a object allocation $x = \kwcl \kwcr$,
a field access $x = x \kwsl x \kwsr$,
a field assignment $x \kwsl x \kwsr = x$,
a conditional branch $\kwif \; x \; \lab \; \lab$,
or a return instruction $\kwret \; x$.

@@

@@

    \text{Primive Values} & \pval &\in& \pvalset &=&
    \strset \uplus \cdots\\

    \text{JavaScript ASTs} & \tree &\in& \treeset\\

    \text{Objects} & \obj &\in& \objset &=&
    \strset \finmap \valset\\
  \end{array}
\]

\todo
% States $\stset$ consist of heaps $\heapset$ and contexts $\ctxtset$.  A heap
% $\heap \in \heapset$ is a finite mapping from addresses to objects.  Each object
% allocation instruction $x = \kwcl \kwcr$ creates a unique address $\addr \in
% \addrset$ different from existing addresses.  A context $\ctxt \in \ctxtset$ is
% a triple of a label, a environment, and an optional calling context $\ctxtopt
% \in \ctxtoptset = \ctxtset \uplus \{ \bot \}$.  An environment $\env \in
% \envset$ is a finite mapping from variables to values. A value $\val \in
% \valset$ is a primitive value $\pval \in \pvalset$ (e.g., a string value $\str
% \in \strset$), an address $\addr \in \addrset$, a JavaScript AST $\tree \in
% \treeset$, or a function $f \in \funcset$.  A object $\obj \in \objset$ is a
% finite mappings from strings to values.

\subsubsection{Concrete Semantics of Functions}

\[
  \framebox{$\sem{\func}: \stset \times \valset^* \rightarrow \stset \times
  \valset$}
\]

\subsubsection{Concrete Semantics of Instructions}

\[
  \framebox{$\sem{\inst}: \stset \rightarrow \stset$}
\]

\todo
% \[
%   \begin{array}{l}
%     \sem{x = \pval}(\heap, (\lab, \env, \ctxtopt)) =\\
%     \qquad (\heap, (\lab, \env[x \mapsto \pval], \ctxtopt)) \\
%     \\
% 
%     \sem{x = \op \kwrl x_1, \cdots, x_n \kwrr}(\heap, (\lab, \env, \ctxtopt)) =\\
%     \qquad (\heap, (\lab, \env[x \mapsto \sem{\op}(x_1, \cdots, x_n)], \ctxtopt)) \\
%     \\
% 
%     \sem{x = x \kwrl x^* \kwrr}(\heap, (\lab, \env, \ctxtopt)) =\\
%     \qquad \todo \\
%     \\
% 
%     \sem{x = \kwcl \kwcr}(\heap, (\lab, \env, \ctxtopt)) =\\
%     \qquad \todo \\
%     \\
% 
%     \sem{x = x \kwsl x \kwsr}(\heap, (\lab, \env, \ctxtopt)) =\\
%     \qquad \todo \\
%     \\
% 
%     \sem{x \kwsl x \kwsr = x}(\heap, (\lab, \env, \ctxtopt)) =\\
%     \qquad \todo \\
%     \\
% 
%     \sem{\kwif \; x \; \lab \; \lab}(\heap, (\lab, \env, \ctxtopt)) =\\
%     \qquad \todo \\
%     \\
% 
%     \sem{\kwret \; x}(\heap, (\lab, \env, \ctxtopt)) =\\
%     \qquad \todo \\
% 
%   \end{array}
% \]

\subsubsection{Concrete Semantics of Expressions}

\[
  \framebox{$\sem{\expr}: \stset \rightarrow \stset \times \valset$}
\]

\subsubsection{Concrete Semantics of References}

\[
  \framebox{$\sem{\refer}: \stset \rightarrow \stset \times \valset$}
\]

@@

A function specialization is defined with the following three components:
\begin{itemize}
  \item Restriction $\restrictset$
  \item Concretization $\gamma: \restrictset \rightarrow \powerset{\stset
    \times \valset^*}$
  \item Transformation $\trans: \restrictset \rightarrow \funcset \rightarrow
    \funcset$.
\end{itemize}

@@

\subsubsection{Constant Propagation Specialization}
\begin{itemize}
  \item Restriction $\cprestrictset = (\valoptset)^* =
    (\valset \uplus \{ \top \})^*$
  \item Concretization $\cpgamma: (\valoptset)^* \rightarrow \powerset{\stset
    \times \valset^*}$ s.t.
    \[
      \cpgamma(\valopt_1, \cdots, \valopt_n) = \stset \times (V_1 \times \cdots
      \times V_n)
    \]
    where $V_i = \left\{
      \begin{array}{ll}
        \valset & \text{if} \; \valopt_i = \top\\
        \{ \valopt_i \} & \text{otherwise}\\
      \end{array}
    \right.$
  \item Transformation $\cptrans: (\valoptset)^* \rightarrow \funcset
    \rightarrow \funcset$
\end{itemize}
In the constant propagation specialization, a restriction is a sequences of
static values $\valset$ or a dynamic value $\top$.

@@

\[
  \framebox{$\cptranse: \exprset \times \senvset \rightarrow \svalset$}
\]

\[
  \begin{array}{lcl}
    \cptranse(\pval, \env) &=& \pval\\
    \cptranse(\op \kwrl \expr^* \kwrr, \env) &=& \todo\\
    \cptranse(\kwcl \kwcr, \env) &=& \todo\\
    \cptranse(\refer, \env) &=& \todo\\
  \end{array}
\]

\[
  \framebox{$\cptransi: \instset \times \senvset \rightarrow \instset \times
  \senvset$}
\]

@@

\[
  \framebox{$\cptrans: (\valoptset)^* \rightarrow \funcset \rightarrow
  \funcset$}
\]

@@

% \subsubsection{Redundancy Elimination Transformation}
% 
% \[
%   \begin{array}{lr@{~}c@{~}l}
%     \text{Restrictions} &\{ \top \}\\
%     \text{Concretization} & \gamma(\top) &=& \stset \times \valset^*\\
%   \end{array}
% \]
% 
% The redundancy elimination transformation is defined without any restriction,
% which means its restriction set consists of a single element $\top$
% representing all possible cases.  Therefore, we omit the restriction part in
% this transformation for the brevity.
% 
% \[
%   \framebox{$\transref: \funcset \rightarrow \funcset$}
% \]
% 
% \todo

>>>>

\[
  \begin{array}{lr@{~}c@{~}c@{~}c@{~}l}
    \text{Functions} & \funcset &\ni& \func &::=&
    \kwdef \; \varx \; \kwrl \varx^* \kwrr \; \lab\\

    \text{Variables} & \varset &\ni& \varx\\

@@

    \text{Instructions} & \instset &\ni& \inst &::=&
    \refer = \expr \mid
    \varx = \expr \kwrl \expr^* \kwrr \mid
    \kwif \; \expr \; \lab \; \lab \mid
    \kwret \; \expr\\

@@

    \text{References} & \referset &\ni& \refer &::=&
    \varx \mid \refer \kwsl \expr \kwsr\\
  \end{array}
\]

An $\ires$ program $\prog = (\istset, \getinst, \getnext)$ consists of initial
states and two mappings; $\getinst: \labset \rightarrow \instset$ maps labels to
their instructions, and $\getnext: \labset \rightarrow \labset$ maps labels to
their next labels, where a label $\lab \in \labset$ denotes a program point.  A
function $\func \in \funcset$ is defined with its name, parameters, and body
label.  For presentation brevity, we assume that no global variables exist in
this paper.  An instruction $\inst \in \instset$ is an assignment $\refer =
\expr$, a function call $x = \expr \kwrl \expr^* \kwrr$, a branch $\kwif \;
\expr \; \lab \; \lab$, or a return instruction $\kwret \; \expr$.  An
invocation of an abstract algorithm in ECMAScript is compiled to a function call
instruction with a new temporary variable.  We represent loops using branch
instructions with cyclic pointing of labels in $\getnext$.  An expression is a
primitive value $\pval$, an operation $\op \kwrl \expr^* \kwrr$, an object
allocation $\kwcl \kwcr$, or a reference $\refer$.  A reference is a variable
$x$ or an object field $\refer \kwsl \expr \kwsr$.  We write $\refer.\varf$ to
briefly represent $\refer \kwsl \code{"f"} \kwsr$.

@@

    \text{Objects} & \obj &\in& \objset &=&
    \strset \finmap \valset\\

@@

    \text{Primive Values} & \pval &\in& \pvalset &=&
    \boolset \uplus \strset \uplus \cdots\\

    \text{JavaScript ASTs} & \tree &\in& \treeset\\
  \end{array}
\]

States $\stset$ consist of labels $\labset$, calling context stacks
$\ctxtset^*$, heaps $\heapset$, and environments $\envset$.  A calling context
$\ctxt \in \ctxtset$ consists of a label denoting the return point, a caller's
environment, and a return variable.  A heap $\heap \in \heapset$ is a finite
mapping from addresses to objects, and an object $\obj \in \objset$ is a finite
mappings from strings to values.  Each object allocation $\kwcl \kwcr$ creates a
unique address $\addr \in \addrset$ different from existing addresses.  An
environment $\env \in \envset$ is a finite mapping from variables to values. A
value $\val \in \valset$ is a primitive value $\pval \in \pvalset$ (e.g., a
boolean value $\bool \in \boolset$ or a string value $\str \in \strset$), an
address $\addr \in \addrset$, a JavaScript AST $\tree \in \treeset$, or a
function $f \in \funcset$.

\subsubsection{Concrete Semantics}

The concrete semantics $\sem{\prog}$ of an $\ires$ program $\prog = (\istset, \getinst,
\getnext)$ is defined as follows:
\[
  \sem{\prog} = \{ \st \in \stset \mid \ist \in \istset \wedge \ist \trans^* \st \}
\]
where $\trans^*$ denotes one or more repetition of $\trans$, and $\st \trans
\st$ if and only if $\st = (\lab, \_, \_, \_)$ and $\sem{\getinst(\lab)}(\st) =
\st'$. Now, we define the denotational semantics of instructions $\sem{\inst}:
\stset \rightarrow \stset$, expressions $\sem{\expr}: \stset \rightarrow \stset
\times \valset$, and references $\sem{\refer}: \stset \rightarrow \stset \times
\valset$.

\paragraph{Instructions: $\framebox{$\sem{\inst}: \stset \rightarrow \stset$}$}
\begin{itemize}
  \item \underline{Variable Assignments}:
    \[
      \sem{\varx = \expr}(\st) =
      (\getnext(\lab), \ctxts, \heap, \env[\varx \mapsto \val])
    \]
    where
    \[
      \sem{\expr}(\st) = ((\lab, \ctxts, \heap, \env), \val)
    \]

  \item \underline{Field Assignments}:
    \[
      \sem{\refer \kwsl \expr_0 \kwsr = \expr_1}(\st) =
      (\getnext(\lab), \ctxts, \heap[\addr \mapsto \obj'], \env)
    \]
    where
    \[
      \begin{array}{l@{~}c@{~}ll}
        \sem{\refer}(\st) &=& (\st', \addr) &\wedge\\
        \sem{\expr_0}(\st') &=& (\st'', \str) &\wedge\\
        \sem{\expr_1}(\st'') &=& ((\lab, \ctxts, \heap, \env), \val) &\wedge\\
        \obj &=& \heap(\addr) &\wedge\\
        \obj' &=& \obj[\str \mapsto \val]\\
      \end{array}
    \]

  \item \underline{Function Calls}:
    \[
      \sem{\varx = \expr \kwrl \expr_1 \cdots \expr_n \kwrr}(\st) =
      (\lab_\varf, \ctxt :: \ctxts, \heap, \env')
    \]
    where
    \[
      \begin{array}{l@{~}c@{~}ll}
        \sem{\expr} &=& (\st_0, \kwdef \; \varf \; \kwrl \varp_1, \cdots,
        \varp_n \kwrr \; \lab_\varf) &\wedge\\
        \sem{\expr_k}(\st_{k-1}) &=& (\st_k, \val_k) \; \forall 1 \leq k \leq n
        &\wedge\\
        \st_n &=& (\lab, \ctxts, \heap, \env) &\wedge\\
        \env' &=& [\varp_1 \mapsto \val_1, \cdots, \varp_n \mapsto \val_n]
        &\wedge\\
        \ctxt &=& (\getnext(\lab), \env, \varx)\\
      \end{array}
    \]

  \item \underline{Branches}:
    \[
      \sem{\kwif \; \expr \; \lab_\vart \; \lab_\varf}(\st) =
      \left\{
        \begin{array}{ll}
          (\lab_\vart, \ctxts, \heap, \env) & \text{if} \; \val = \true\\
          (\lab_\varf, \ctxts, \heap, \env) & \text{if} \; \val = \false\\
        \end{array}
      \right.
    \]
    where
    \[
      \sem{\expr}(\st) = ((\lab, \ctxts, \heap, \env), \val)
    \]

  \item \underline{Returns}:
    \[
      \sem{\kwret \; \expr}(\st) = (\lab, \ctxts, \heap, \env[\varx \mapsto
      \val])
    \]
    where
    \[
      \sem{\expr}(\st) = ((\_, (\lab, \env, \varx) :: \ctxts, \heap, \_), \val)
    \]
\end{itemize}

\paragraph{Expressions: $\framebox{$\sem{\expr}: \stset \rightarrow \stset
\times \valset$}$}
\begin{itemize}
  \item \underline{Primive Values}:
    \[
      \sem{\pval}(\st) = \todo
    \]
  \item \underline{Operations}:
    \[
      \sem{\op \kwrl \expr^* \kwrr}(\st) = \todo
    \]
  \item \underline{Object Allocations}:
    \[
      \sem{\kwcl \kwcr}(\st) = \todo
    \]
  \item \underline{References}:
    \[
      \sem{\refer}(\st) = \todo
    \]
\end{itemize}

\paragraph{References: $\framebox{$\sem{\refer}: \stset \rightarrow \stset
\times \valset$}$}
\begin{itemize}
  \item \underline{Variable Lookups}:
    \[
      \sem{\varx}(\st) = \todo
    \]
  \item \underline{Field Lookups}:
    \[
      \sem{\refer \kwsl \expr \kwsr}(\st) = \todo
    \]
\end{itemize}

@@

A function specialization is defined with the following three components:
\begin{itemize}
  \item Restriction $\restrictset$
  \item Concretization $\gamma: \restrictset \rightarrow \powerset{\stset
    \times \valset^*}$
  \item Transformation $\transform: \restrictset \rightarrow \funcset
    \rightarrow \funcset$.
\end{itemize}

@@

\subsubsection{Constant Propagation Specialization}
\begin{itemize}
  \item Restriction $\cprestrictset = (\valoptset)^* =
    (\valset \uplus \{ \top \})^*$
  \item Concretization $\cpgamma: (\valoptset)^* \rightarrow \powerset{\stset
    \times \valset^*}$ s.t.
    \[
      \cpgamma(\valopt_1, \cdots, \valopt_n) = \stset \times (V_1 \times \cdots
      \times V_n)
    \]
    where $V_i = \left\{
      \begin{array}{ll}
        \valset & \text{if} \; \valopt_i = \top\\
        \{ \valopt_i \} & \text{otherwise}\\
      \end{array}
    \right.$
  \item Transformation $\cptransform: (\valoptset)^* \rightarrow \funcset
    \rightarrow \funcset$
\end{itemize}
In the constant propagation specialization, a restriction is a sequences of
static values $\valset$ or a dynamic value $\top$.

@@

\[
  \framebox{$\cptransforme: \exprset \times \senvset \rightarrow \svalset$}
\]

\[
  \begin{array}{lcl}
    \cptransforme(\pval, \env) &=& \pval\\
    \cptransforme(\op \kwrl \expr^* \kwrr, \env) &=& \todo\\
    \cptransforme(\kwcl \kwcr, \env) &=& \todo\\
    \cptransforme(\refer, \env) &=& \todo\\
  \end{array}
\]

\[
  \framebox{$\cptransformi: \instset \times \senvset \rightarrow \instset \times
  \senvset$}
\]

@@

\[
  \framebox{$\cptransform: (\valoptset)^* \rightarrow \funcset \rightarrow
  \funcset$}
\]

@@

% \subsubsection{Redundancy Elimination Transformation}
% 
% \[
%   \begin{array}{lr@{~}c@{~}l}
%     \text{Restrictions} &\{ \top \}\\
%     \text{Concretization} & \gamma(\top) &=& \stset \times \valset^*\\
%   \end{array}
% \]
% 
% The redundancy elimination transformation is defined without any restriction,
% which means its restriction set consists of a single element $\top$
% representing all possible cases.  Therefore, we omit the restriction part in
% this transformation for the brevity.
% 
% \[
%   \framebox{$\transformref: \funcset \rightarrow \funcset$}
% \]
% 
% \todo
